---
title: "ENV 797 - Time Series Analysis for Energy and Environment Applications | Spring 2025"
subtitle: "Assignment 7 - Due date 03/06/25"
author: "Ayoung Kim"
output: pdf_document
geometry: margin=2.54cm
---

## Directions

You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. And to do so you will need to fork our repository and link it to your RStudio.  

Once you have the file open on your local machine the first thing you will do is rename the file such that it includes your first and last name (e.g., "LuanaLima_TSA_A07_Sp25.Rmd"). Then change "Student Name" on line 4 with your name. 

Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Submit this pdf using Sakai. 

Packages needed for this assignment: "forecast","tseries". Do not forget to load them before running your script, since they are NOT default packages.\\

## Set up

```{r, Warning=FALSE}
#Load/install required package here
library(forecast)
library(tseries)
library(dplyr)
library(ggplot2)
```

## Importing and processing the data set

Consider the data from the file "Net_generation_United_States_all_sectors_monthly.csv". The data corresponds to the monthly net generation from January 2001 to December 2020 by source and is provided by the US Energy Information and Administration. **You will work with the natural gas column only**.


### Q1

Import the csv file and create a time series object for natural gas. Make you sure you specify the **start=** and **frequency=** arguments. Plot the time series over time, ACF and PACF.

```{r}
library(openxlsx)
library(readxl)

#Importing the csv file 
net <- read.csv(file="./Net_generation_United_States_all_sectors_monthly.csv",skip=4)

#Checking the colnames 
colnames(net)

#Rename the column 
net_generation_rename <- net %>% rename(
  Natural_gas = `natural.gas.thousand.megawatthours`
)

#Selected only Month, Natural gas 
net_generation <- select(net_generation_rename,Natural_gas) #Month

#Reversed sequence of values from old to recent one
net_generation_reversed <- net_generation %>%
  mutate(
    Natural_gas = rev(Natural_gas)
  )

#Checking the order of values and whether it matches with Year-Month
head(net_generation_reversed)

#Create a time series object 
ts_net_generation<-ts(net_generation_reversed,start=c(2000,1),frequency=12)


#Plot the time series (I checked my code with AI to get rid of the Month)
autoplot(ts_net_generation)

acf_net_generation <- Acf(ts_net_generation, lag.max = 60, plot = FALSE)
autoplot(acf_net_generation)

pacf_net_generation <- Pacf(ts_net_generation, lag.max = 60, plot = FALSE)
autoplot(pacf_net_generation)



```

### Q2

Using the $decompose()$ and the $seasadj()$ functions create a series without the seasonal component, i.e., a deseasonalized natural gas series. Plot the deseasonalized series over time and corresponding ACF and PACF. Compare with the plots obtained in Q1.

>Answer: Compared to the plots in Q1, the plots of the deseasoned time series and the corresponding ACF/PACF show a reduced seasonal component. This is particularly evident in the 'deseasonalized plot with the original curve,' where the deseasoned curve is compared to the original. Additionally, in the ACF and PACF plots, the seasonal components have been removed.

```{r}
#Deseasonalizing with decompose() function and its plot 
decompose_net_generation<-decompose(ts_net_generation)
plot(decompose_net_generation)

#Deseasonalizing with seasadj() and its plot 
deseasonal_net_generation<-seasadj(decompose_net_generation)

#Deseasonalized plot with the original time series 
autoplot(ts_net_generation,series="Original")+
  autolayer(deseasonal_net_generation,series="Deseasoned")

#Corresponding ACF and PACF 
acf_deseasonal_net_generation <- Acf(deseasonal_net_generation,lag=60,plot = FALSE)
plot(acf_deseasonal_net_generation, main="ACF of deseasonalized net generation of natural gas")

pacf_deseasonal_net_generation <- Pacf(deseasonal_net_generation,lag=60,plot = FALSE)
plot(pacf_deseasonal_net_generation, main="PACF of deseasonalized net generation of natural gas")

```

## Modeling the seasonally adjusted or deseasonalized series

### Q3

Run the ADF test and Mann Kendall test on the deseasonalized data from Q2. Report and explain the results.
>Answer: According to the ADF test, the lag order is 6, the Dickey-Fuller value is -4.0271, and the p-value is 0.01. The null hypothesis would be rejected due to statistical significance and the series would be stationary. In the Mann-Kendall test result, the p-value is ≤ 2.22e-16, indicating a significant trend.

```{r, Warning=FALSE}
#Installing ManKendall 
library(Kendall)

#Change the format of deseasonsal_net_generation to numeric 
numeric_deseasonal_net_generation <-as.numeric(deseasonal_net_generation)

#ADF test 
adf_test_1 <-adf.test(numeric_deseasonal_net_generation)
print(adf_test_1)

#Man Kendall Test 
Mann_Kendall_1 <-MannKendall(numeric_deseasonal_net_generation)
summary(Mann_Kendall_1)
```
### Q4
Using the plots from Q2 and test results from Q3 identify the ARIMA model parameters $p,d$ and $q$. Note that in this case because you removed the seasonal component prior to identifying the model you don't need to worry about seasonal component. Clearly state your criteria and any additional function in R you might use. DO NOT use the $auto.arima()$ function. You will be evaluated on ability to understand the ACF/PACF plots and interpret the test results.

>Answer: 
ARIMA model parameter (p,d,q)
p=1, because the pacf shows the cut off after lag 1, likely to be AR(1) model
d=0, because it was deseasonalized already. 
q=0, because ACF plot shows the slow decay, which means it is AR model. 
Therefore, the ARIMA model would be ARIMA (1,0,0). According to the summary of arima_net_generation, the coefficient of AR(1) would be 0.9825 and it means that the previous value would have significant impacts on forecasting the next value. 


```{r}
#ARIMA model with parameters 
arima_net_generation <- arima(deseasonal_net_generation, order = c(1, 0, 0), include.mean = TRUE)
summary(arima_net_generation)
```

### Q5

Use `Arima()` from package "forecast" to fit an ARIMA model to your series considering the order estimated in Q4. You should allow constants in the model, i.e., `include.mean = TRUE` or `include.drift=TRUE`. **Print the coefficients** in your report. Hint: use the `cat()` or `print()` function to print.

```{r}
arima_net_generation1 <- Arima(deseasonal_net_generation, order = c(1, 0, 0),include.mean = TRUE)
print(arima_net_generation1)

#Plot the residuals of the ARIMA along with residuals ACF and PACF
checkresiduals(arima_net_generation1)

```

### Q6

Now plot the residuals of the ARIMA fit from Q5 along with residuals ACF and PACF on the same window. You may use the $checkresiduals()$ function to automatically generate the three plots. Do the residual series look like a white noise series? Why?

>Answer: The residueal series looks like a white noise series with random noises around 0. 
(Check the answer later)

## Modeling the original series (with seasonality)

### Q7

Repeat Q3-Q6 for the original series (the complete series that has the seasonal component). Note that when you model the seasonal series, you need to specify the seasonal part of the ARIMA model as well, i.e., $P$, $D$ and $Q$.

```{r}
#Installing ManKendall 
library(Kendall)

#Change the format of deseasonsal_net_generation to numeric 
numeric_deseasonal_net_generation <-as.numeric(deseasonal_net_generation)

#ADF test 
adf_test_1 <-adf.test(numeric_deseasonal_net_generation)
print(adf_test_1)

#Man Kendall Test 
Mann_Kendall_1 <-MannKendall(numeric_deseasonal_net_generation)
summary(Mann_Kendall_1)


```


### Q8

Compare the residual series for Q7 and Q6. Can you tell which ARIMA model is better representing the Natural Gas Series? Is that a fair comparison? Explain your response.

```{r}

```


## Checking your model with the auto.arima()

**Please** do not change your answers for Q4 and Q7 after you ran the $auto.arima()$. It is **ok** if you didn't get all orders correctly. You will not loose points for not having the same order as the $auto.arima()$.

### Q9

Use the $auto.arima()$ command on the **deseasonalized series** to let R choose the model parameter for you. What's the order of the best ARIMA model? Does it match what you specified in Q4?

```{r}

```


### Q10

Use the $auto.arima()$ command on the **original series** to let R choose the model parameters for you. Does it match what you specified in Q7?

```{r}

```

