---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2025"
subtitle: "Assignment 2 - Ayoung Kim "
author: "Ayoung Kim"
output: pdf_document
geometry: margin=2.54cm
---

## Submission Instructions

You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github.   

Once you have the file open on your local machine the first thing you will do is rename the file such that it includes your first and last name (e.g., "LuanaLima_TSA_A02_Sp24.Rmd"). Then change "Student Name" on line 4 with your name. 

Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Submit this pdf using Sakai. 

## R packages

R packages needed for this assignment:"forecast","tseries", and "dplyr". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r, include=FALSE}
setwd("/Users/ayoungkim/TSA_Sp25")
getwd()

```



```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=60), tidy=TRUE) 
#the tidy arguments prevent the code from running off the page i.e. wrap by itself
```


```{r}
#Load/install required package here
#install.packages("forecast")
#install.packages("tseries")
#install.packages("dplyr")

#Addtional Package (For Plot)
#install.packages("ggplot2")
#install.packages("formatR")

library(forecast)
library(tseries)
library(dplyr)
library(ggplot2)
```

## Data set information  

Consider the data provided in the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx" on our **Data** folder. The data comes from the US Energy Information and Administration and corresponds to the December 2023 Monthly Energy Review. The spreadsheet is ready to be used. You will also find a $.csv$ version of the data "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source-Edit.csv". You may use the function $read.table()$ to import the $.csv$ data in R. Or refer to the file "M2_ImportingData_CSV_XLSX.Rmd" in our Lessons folder for functions that are better suited for importing the $.xlsx$.

```{r package, message=FALSE}
#Loading packages 
#install.packages("readxl")
#install.packages("openxlsx")

library(readxl)
library(openxlsx)
```

```{r}

#Setting a working directory again 
setwd("/Users/ayoungkim/TSA_Sp25")

renewable_data2 <- read.xlsx(xlsxFile="./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",sheet = "Monthly Data", startRow = 13, colNames = FALSE)

read_col_names2  <- read.xlsx(xlsxFile="./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",sheet = "Monthly Data", rows = 11, colNames = FALSE)

#Assign the column names to the data set
colnames(renewable_data2) <- read_col_names2

#Visualize the first rows of the data set
head(renewable_data2)

```


## Question 1
You will work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series only. Use the command head() to verify your data.

```{r}
#Using "select" function, selected only Total Biomass Energy Production, Total Renewable Energy Production, and Hydroelectric Powr Consumption. 
renewable_data2_filtered <- select(renewable_data2,`Total Biomass Energy Production`,`Total Renewable Energy Production`,`Hydroelectric Power Consumption`)

df_renewable_data2_filtered<-as.data.frame(renewable_data2_filtered)

head(df_renewable_data2_filtered)
```


## Question 2
Transform your data frame in a time series object and specify the starting point and frequency of the time series using the function ts().

```{r}
#Transform into a time series object + Starting point and frequency 
#1 Time Series of Total Biomass Energy Production 
ts1_renewable_data2_filtered <- ts(df_renewable_data2_filtered$`Total Biomass Energy Production`,start = c(1,1),frequency = 12)

ts1_renewable_data2_filtered

#2 Time series of Total Renewable Energy Production 
ts2_renewable_data2_filtered<-ts(df_renewable_data2_filtered$`Total Renewable Energy Production`, start = c(1,1), frequency=12)

ts2_renewable_data2_filtered

#3 Time series of Hydroelectricc Power Consumption  
ts3_renewable_data2_filtered <-ts(df_renewable_data2_filtered$`Hydroelectric Power Consumption`,start = c(1,1), frequency = 12)

ts3_renewable_data2_filtered

```


## Question 3
Compute mean and standard deviation for these three series.

```{r}
#1 Mean and standard deviation of Time Series of Total Biomass Energy Production 
mean(ts1_renewable_data2_filtered)
sd(ts1_renewable_data2_filtered)

#2 Mean and standard deviation of Time Series of Total Renewable Enerfy Production 
mean(ts2_renewable_data2_filtered)
sd(ts2_renewable_data2_filtered)

#3 Mean and standard deviation of Time Series of Hydroelectric Power Consumption 
mean(ts3_renewable_data2_filtered)
sd(ts3_renewable_data2_filtered)

```


## Question 4
Display and interpret the time series plot for each of these variables. Try to make your plot as informative as possible by writing titles, labels, etc. For each plot add a horizontal line at the mean of each series in a different color.

Plot 1, which represents the time series of total biomass energy production, shows the increasing trend. The mean is around 282 with the standard deviation of 94.1. It has high inflow because the time series gradually increase. 

Plot 2 displaying time series of total renewable energy produciton shows the increasing trend as well. The mean is around 402 with the standard deviation of 143.9. It has high inflow because the time series gradually increase. 

In the Plot 3 representing the time series of hydroelectric power consumption, I could find the seasonal pattern. The mean is around 79 and it has lower standard deviation with 14.11 than other 2 time series, which are 94.1 and 143. 9 each. IT has low inflow because the time series fluctuate with certain pattern. 

Checked my code using AI. 


```{r}
## Titles, labels, horizontal line at the mean 
## Plot, interpretation 

#Plot 1 - Time series of total biomass energy production 
autoplot(ts1_renewable_data2_filtered) +
  labs(title = "Time Series of Total Biomass Energy Production",  
       y = "Total Biomass Energy Production",  
       x = "Time") + 
  geom_hline(yintercept = mean(ts1_renewable_data2_filtered, na.rm = TRUE), 
             color = "blue", linetype = "solid", size = 1) 

#Plot 2 - Time Series of Total Renewable Energy Production
autoplot(ts2_renewable_data2_filtered) +
  labs(title = "Time Series of Total Renewable Energy Production",  
       y = "Total Renewable Energy Production",  
       x = "Time") + 
  geom_hline(yintercept = mean(ts2_renewable_data2_filtered, na.rm = TRUE), 
             color = "red", linetype = "solid", size = 1) 

#Plot 3 - Time Series of Hydroelectric Power Consumption 
autoplot(ts3_renewable_data2_filtered) +
  labs(title = "Time Series of Hydroelectric Power Consumption",  
       y = "Hydroelectric Power Consumption",  
       x = "Time") + 
  geom_hline(yintercept = mean(ts3_renewable_data2_filtered, na.rm = TRUE), 
             color = "purple", linetype = "solid", size = 1) 

```


## Question 5
Compute the correlation between these three series. Are they significantly correlated? Explain your answer.

ts1 - Total Biomass Energy Production
ts2 - Total Renewable Energy Production 
ts3 - Hydroelectric Power Consumption 

Answer: ts1 and ts2 are significantly correlated seeing that the result of correlation computation is 0.9678. However, ts2 & ts3 and ts 1 &3 are not correlated significantly wit the result of -0.02916103 and -0.1142927 each. 


```{r}
#Making as one df to find the correlation between three timeseries in matrix 
df_ts_correlation <-data.frame(ts1_renewable_data2_filtered,ts2_renewable_data2_filtered,ts3_renewable_data2_filtered)

correlation_ts<-cor(df_ts_correlation)
correlation_ts

#OR 

#1 ts1+ts2 
cor(ts1_renewable_data2_filtered,ts2_renewable_data2_filtered)

#2 ts2+ts3
cor(ts2_renewable_data2_filtered,ts3_renewable_data2_filtered)

# ts1+ts3
cor(ts1_renewable_data2_filtered,ts3_renewable_data2_filtered)

```


## Question 6
Compute the autocorrelation function from lag 1 up to lag 40 for these three variables. What can you say about these plots? Do the three of them have the same behavior?

Answer: The autocorrelation of ts1 and ts2 shows a similar behavior with a downward trend. The ACF values decrease as the lag increases in both cases. However, the autocorrelation of ts3 (Hydroelectric Power Consumption) exhibits a different behavior. It shows a seasonal pattern with regular fluctuations, rising and falling as the lag increases. 


##Each of the bar represents of the p
```{r}
#1 Autocorrelation of ts1 
acf_ts1 <-acf(ts1_renewable_data2_filtered,lag.max = 40,main="Autocorrelation of Total Biomass Energy Production")

#2 Autocorrelation of ts2 
acf_ts2<-acf(ts2_renewable_data2_filtered,lag.max = 40, main="Autocorrelation of Total Renewble Energy Production")

#3 Autocorrelation of ts3 
acf_ts3<-acf(ts3_renewable_data2_filtered,lag.max = 40, main="Autrocorrelation of Hydroelectric Power Consumption")

```


## Question 7
Compute the partial autocorrelation function from lag 1 to lag 40 for these three variables. How these plots differ from the ones in Q6?

Answer: In Q6, the autocorrelation (ACF) values for ts1 and ts2 were all positive, but in the partial autocorrelation (PACF) plots, the values turned out to be negative. The PACF plots in Q7 show the direct correlations, excluding the influence of previous lags. The PACF plot for Hydroelectric Power Consumption in Q7 has a similar pattern to the one in Q6, but the distribution is less fluctuating compared to Q6.

```{r}
#1 Partial Autocorrelation of ts1 
pacf_ts1 <-pacf(ts1_renewable_data2_filtered,lag.max = 40,main="Partial Autocorrelation of Total Biomass Energy Production")

#2 Partial Autocorrelation of ts2 
pacf_ts2<-pacf(ts2_renewable_data2_filtered,lag.max = 40, main="Partial Autocorrelation of Total Renewble Energy Production")

#3 Partial Autocorrelation of ts3 
pacf_ts3<-pacf(ts3_renewable_data2_filtered,lag.max = 40, main="Partial Autocorrelation of Hydroelectric Power Consumption")
```

