---
title: "ENV 797 - Time Series Analysis for Energy and Environment Applications | Spring 2025"
subtitle: "Assignment 6 - Due date 02/27/25"
author: "Ayoung Kim"
output: pdf_document
geometry: margin=2.54cm
---

## Directions

You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github.

Once you have the file open on your local machine the first thing you will do is rename the file such that it includes your first and last name (e.g., "LuanaLima_TSA_A06_Sp25.Rmd"). Then change "Student Name" on line 4 with your name.

Then you will start working through the assignment by **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Submit this pdf using Sakai.

R packages needed for this assignment: "ggplot2", "forecast", "tseries" and "sarima". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.

```{r, Warning=FALSE}
#Load/install required package here
install.packages("ggplot2")
install.packages("forecast")
install.packages("tseries")
install.packages("sarima")

library(ggplot2)
library(forecast)
library(tseries)
library(sarima)


```

This assignment has general questions about ARIMA Models.

## Q1

Describe the important characteristics of the sample autocorrelation function (ACF) plot and the partial sample autocorrelation function (PACF) plot for the following models:

* AR(2)

> Answer: In the AR model, ACF will decay exponentially with time, which means that there is a relationship with a previous lag. PACF will identify the order of the AR model. The PACF will cut off at the order of the AR model, which is 2 in AR(2). 


* MA(1)

> Answer: Different from AR, in MA models, ACF will identify the order of the MA model and the PACF will decay exponentially. In MA(1), ACF will show the cut off at 1 without other significant lags after that.  

## Q2

Recall that the non-seasonal ARIMA is described by three parameters ARIMA$(p,d,q)$ where $p$ is the order of the autoregressive component, $d$ is the number of times the series need to be differenced to obtain stationarity and $q$ is the order of the moving average component. If we don't need to difference the series, we don't need to specify the "I" part and we can use the short version, i.e., the ARMA$(p,q)$.

(a) Consider three models: ARMA(1,0), ARMA(0,1) and ARMA(1,1) with parameters $\phi=0.6$ and $\theta= 0.9$. The $\phi$ refers to the AR coefficient and the $\theta$ refers to the MA coefficient. Use the `arima.sim()` function in R to generate $n=100$ observations from each of these three models. Then, using `autoplot()` plot the generated series in three separate graphs.

Q1. Using random seed? 
Q2. Making it to TS and then autoplot? 

```{r}
#Is it using random seed?
set.seed(123)

arma10<-arima.sim(n=100, model=list(ar=0.6))
arma01<-arima.sim(n=100, model=list(ma=0.9))
arma11<-arima.sim(n=100, model = list(ar=0.6, ma=0.9))

#Plotting generated series in three seperate graphs 

#1 ARIMA (1,0)
plot_arma10<-autoplot(arma10)
plot_arma10

#2 ARIMA (0,1)
plot_arma01<-autoplot(arma01)
plot_arma01

#3 ARIMA (1,1)
plot_arma11<-autoplot(arma11)
plot_arma11

```

(b) Plot the sample ACF for each of these models in one window to facilitate comparison (Hint: use `cowplot::plot_grid()`).


```{r}
#Make ACF plots of each model 
acf_plot_arma10<-autoplot(Acf(arma10,plot = FALSE)) + ggtitle("ACF of ARMA (1,0)")
acf_plot_arma01<-autoplot(Acf(arma01,plot = FALSE)) + ggtitle("ACF of ARMA (0,1)")
acf_plot_arma11<-autoplot(Acf(arma11,plot = FALSE)) + ggtitle("ACF of ARMA (1,1)")

#In one window 
cowplot::plot_grid(acf_plot_arma10,acf_plot_arma01,acf_plot_arma11)


```

(c) Plot the sample PACF for each of these models in one window to facilitate comparison.

```{r}
#Make PACF plots of each model 
pacf_plot_arma10<-autoplot(Pacf(arma10,plot = FALSE)) + ggtitle("PACF of ARMA (1,0)")
pacf_plot_arma01<-autoplot(Pacf(arma01,plot = FALSE)) + ggtitle("PACF of ARMA (0,1)")
pacf_plot_arma11<-autoplot(Pacf(arma11,plot = FALSE)) + ggtitle("PACF of ARMA (1,1)")

#In one window 
cowplot::plot_grid(pacf_plot_arma10,pacf_plot_arma01,pacf_plot_arma11)

```

(d) Look at the ACFs and PACFs. Imagine you had these plots for a data set and you were asked to identify the model, i.e., is it AR, MA or ARMA and the order of each component. Would you be able identify them correctly? Explain your answer.

> Answer: Yes, I can identify them with the cut off and whether it lags are decaying exponentially or not. 

(e) Compare the PACF values R computed with the values you provided for the lag 1 correlation coefficient, i.e., does $\phi=0.6$ match what you see on PACF for ARMA(1,0), and ARMA(1,1)? Should they match?

> Answer:No, they were not matched. 
Q. Should they match? 

(f) Increase number of observations to $n=1000$ and repeat parts (b)-(e).

```{r}
# (a)
arma10_2<-arima.sim(n=1000, model=list(ar=0.6))
arma01_2<-arima.sim(n=1000, model=list(ma=0.9))
arma11_2<-arima.sim(n=1000, model = list(ar=0.6, ma=0.9))

#(b)
#Make ACF plots of each model (n=1000)
acf_plot_arma10_2<-autoplot(Acf(arma10_2,plot = FALSE)) + ggtitle("ACF of ARMA (1,0), n=1000")
acf_plot_arma01_2<-autoplot(Acf(arma01_2,plot = FALSE)) + ggtitle("ACF of ARMA (0,1), n=1000")
acf_plot_arma11_2<-autoplot(Acf(arma11_2,plot = FALSE)) + ggtitle("ACF of ARMA (1,1), n=1000")

#In one window 
cowplot::plot_grid(acf_plot_arma10_2,acf_plot_arma01_2,acf_plot_arma11_2)

#(c)
#Make PACF plots of each model (n=1000)
pacf_plot_arma10_2<-autoplot(Pacf(arma10_2,plot = FALSE)) + ggtitle("PACF of ARMA (1,0), n=1000")
pacf_plot_arma01_2<-autoplot(Pacf(arma01_2,plot = FALSE)) + ggtitle("PACF of ARMA (0,1), n=1000")
pacf_plot_arma11_2<-autoplot(Pacf(arma11_2,plot = FALSE)) + ggtitle("PACF of ARMA (1,1), n=1000")

#In one window 
cowplot::plot_grid(pacf_plot_arma10_2,pacf_plot_arma01_2,pacf_plot_arma11_2)

#(d)Look at the ACFs and PACFs. Imagine you had these plots for a data set and you were asked to identify the model, i.e., is it AR, MA or ARMA and the order of each component. Would you be able identify them correctly? Explain your answer.

#Answer: 

#(e) Compare the PACF values R computed with the values you provided for the lag 1 correlation coefficient, i.e., does $\phi=0.6$ match what you see on PACF for ARMA(1,0), and ARMA(1,1)? Should they match?

#Answer: For PACF, yes, the lag 1 correlation coefficient matched.

```

## Q3

Consider the ARIMA model $y_t=0.7*y_{t-1}-0.25*y_{t-12}+a_t-0.1*a_{t-1}$

(a) Identify the model using the notation ARIMA$(p,d,q)(P,D,Q)_ s$, i.e., identify the integers $p,d,q,P,D,Q,s$ (if possible) from the equation.

p=1
d= How can I identify differencing? 
q=1
P= #Question 
D= How can I identify differencing? 
Q= #Question 
s (seasonal period)= 12

(b) Also from the equation what are the values of the parameters, i.e., model coefficients.

Answer: 
Coefficient of AR = 0.7
Coefficient of MA = -0.1
Coefficient of AR at lag 12 = -0.25

Q. Finding coefficients? 

## Q4

Simulate a seasonal ARIMA$(0, 1)\times(1, 0)_{12}$ model with $\phi =0 .8$ and $\theta = 0.5$ using the `sim_sarima()` function from package `sarima`. The $12$ after the bracket tells you that $s=12$, i.e., the seasonal lag is 12, suggesting monthly data whose behavior is repeated every 12 months. You can generate as many observations as you like. Note the Integrated part was omitted. It means the series do not need differencing, therefore $d=D=0$. Plot the generated
series using `autoplot()`. Does it look seasonal?

```{r}

```

## Q5

Plot ACF and PACF of the simulated series in Q4. Comment if the plots are well representing the model you simulated, i.e., would you be able to identify the order of both non-seasonal and seasonal components from the plots? Explain.

```{r}

```
