---
title: "ENV 797 - Time Series Analysis for Energy and Environment Applications | Spring 2025"
subtitle: "Assignment 5 - Due date 02/18/25"
author: "Ayoung Kim"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## Directions
  
You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. And to do so you will need to fork our repository and link it to your RStudio.  

Once you have the file open on your local machine the first thing you will do is rename the file such that it includes your first and last name (e.g., "LuanaLima_TSA_A05_Sp25.Rmd"). Then change "Student Name" on line 4 with your name. 

Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Submit this pdf using Sakai. 

R packages needed for this assignment: "readxl", "ggplot2", "forecast","tseries", and "Kendall". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here
library(forecast)
library(tseries)
library(ggplot2)
library(Kendall)
library(lubridate)
library(tidyverse)  #load this package so yon clean the data frame using pipes
library(zoo)
library(dplyr)

```

## Decomposing Time Series

Consider the same data you used for A04 from the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx". The data comes from the US Energy Information and Administration and corresponds to the December 2023 Monthly Energy Review.

```{r}
#Importing data set - using xlsx package
library(openxlsx)
library(readxl)

getwd()
setwd("/Users/ayoungkim/TSA_Sp25/Data")
getwd()

renewable_data2 <- read.xlsx(xlsxFile="./Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",
                             sheet = "Monthly Data", 
                             startRow = 13,
                             colNames = FALSE)

read_col_names2  <- read.xlsx(xlsxFile="./Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",
                              sheet = "Monthly Data", 
                              rows = 11, 
                              colNames = FALSE)

colnames(renewable_data2) <- read_col_names2
head(renewable_data2)

nobs=nrow(renewable_data2)
nvar=ncol(renewable_data2)

```

### Q1

For this assignment you will work only with the following columns: Solar Energy Consumption and Wind Energy Consumption. Create a data frame structure with these two time series only and the Date column. Drop the rows with *Not Available* and convert the columns to numeric. You can use filtering to eliminate the initial rows or convert to numeric and then use the drop_na() function. If you are familiar with pipes for data wrangling, try using it!

```{r}
#Using "select" function, selected only Solar Energy Consumption and Wind Energy Consumtpion 
renewable_data2_filtered <- select(renewable_data2, 'Month', 'Solar Energy Consumption', 'Wind Energy Consumption')

start_date <- "1973-01-01"
renewable_data2_filtered$Month <- as.Date(seq(from = as.Date(start_date), 
                                               by = "month", 
                                               length.out = nobs))

head(renewable_data2_filtered$Month)

# Converted the format to Year-Month
renewable_data2_filtered$Month <- format(renewable_data2_filtered$Month, "%Y-%m")
head(renewable_data2_filtered)

#Convert to numeric
renewable_data2_filtered$`Solar Energy Consumption` <- as.numeric(renewable_data2_filtered$`Solar Energy Consumption`)
renewable_data2_filtered$`Wind Energy Consumption` <- as.numeric(renewable_data2_filtered$`Wind Energy Consumption`)

#Dropped 'Not available' 
renewable_data2_filtered <- drop_na(renewable_data2_filtered)
head(renewable_data2_filtered)

#1 ts1 - Solar energy consumption 
ts1_renewable_data2_filtered<-ts(renewable_data2_filtered$'Solar Energy Consumption', 
                                 start = c(1,1), 
                                 frequency=12)
ts1_renewable_data2_filtered

#2 ts2 - Wind Energy Consumption
ts2_renewable_data2_filtered<-ts(renewable_data2_filtered$'Wind Energy Consumption', 
                                 start = c(1,1), 
                                 frequency=12)
ts2_renewable_data2_filtered

#Converting time series to numeric
solar_energy_numeric <- as.numeric(ts1_renewable_data2_filtered)
wind_energy_numeric <- as.numeric(ts2_renewable_data2_filtered)

#Make two ts and month as a dataframe
df_ts_renewable_data2_filtered <- data.frame(
   Month = renewable_data2_filtered$Month,
   Solar_Energy_Consumption = solar_energy_numeric,
   Wind_Energy_Consumption = wind_energy_numeric)

df_ts_renewable_data2_filtered

```

### Q2

Plot the Solar and Wind energy consumption over time using ggplot. Plot each series on a separate graph. No need to add legend. Add informative names to the y axis using `ylab()`. Explore the function `scale_x_date()` on ggplot and see if you can change the x axis to improve your plot. Hint: use `scale_x_date(date_breaks = "5 years", date_labels = "%Y")")`

```{r}
# Convert the 'Month' column to Date format (Confirmed my code with gpt)
df_ts_renewable_data2_filtered$Month <- as.Date(paste0(df_ts_renewable_data2_filtered$Month, "-01"), format = "%Y-%m-%d")

#1 Plot - Solar Energy Consumption + ylab, scale_x_date
plot_ts1 <- ggplot(df_ts_renewable_data2_filtered, aes(x = Month, y = Solar_Energy_Consumption)) + 
  geom_line() +  
  ylab("Solar Energy Consumption") + 
   xlab("Date (Year)")+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y")  

print(plot_ts1)

#2 Plot - Wind Energy Consumption 
plot_ts2 <- ggplot(df_ts_renewable_data2_filtered, 
                   aes(x = Month, 
                       y = Wind_Energy_Consumption)) + 
  ylab("Wind Energy Consumption") + 
   xlab("Date (Year)")+
  geom_line()+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y")  

print(plot_ts2)
```

### Q3

Now plot both series in the same graph, also using ggplot(). Use function `scale_color_manual()` to manually add a legend to ggplot. Make the solar energy consumption red and wind energy consumption blue. Add informative name to the y axis using `ylab("Energy Consumption)`. And use function `scale_x_date()` to set x axis breaks every 5 years.

```{r}
plot_ts_combined <- ggplot() +
  geom_line(data = df_ts_renewable_data2_filtered, 
            aes(x = Month, 
                y = Solar_Energy_Consumption, 
                color = "Solar Energy Consumption")) + 
  geom_line(data = df_ts_renewable_data2_filtered, 
            aes(x = Month,
                y = Wind_Energy_Consumption, 
                color = "Wind Energy Consumption")) +
  scale_color_manual(values = c("Solar Energy Consumption" = "blue", "Wind Energy Consumption" = "green")) +  
  xlab("Date (Year)")+
  ylab("Energy Consumption") +  
  scale_x_date(date_breaks = "5 years", date_labels = "%Y")

print(plot_ts_combined)
```


## Decomposing the time series

The stats package has a function called decompose(). This function only take time series object. As the name says the decompose function will decompose your time series into three components: trend, seasonal and random. This is similar to what we did in the previous script, but in a more automated way. The random component is the time series without seasonal and trend component. 

Additional info on `decompose()`.

1) You have two options: alternative and multiplicative. Multiplicative models exhibit a change in frequency over time.
2) The trend is not a straight line because it uses a moving average method to detect trend. 
3) The seasonal component of the time series is found by subtracting the trend component from the original data then grouping the results by month and averaging them. 
4) The random component, also referred to as the noise component, is composed of all the leftover signal which is not explained by the combination of the trend and seasonal component. 

### Q4

Transform wind and solar series into a time series object and apply the decompose function on them using the additive option, i.e., `decompose(ts_data, type = "additive")`. What can you say about the trend component? What about the random component? Does the random component look random? Or does it appear to still have some seasonality on it?

>Answer: First of all, the trend of solar (ts1) is upwarding. It seems like it still have some seasonality from 0 to 25 lag, while it looks random after that. The trend of wind (ts2) is also upwarding overall, while there is a slight drop at the end. Similar to the former one, it seems like it still have some seasonality from 0 to 27 lag, while it looks random after that. 

```{r}
#Wind and Solar series - Already transformed into a time series object in Q1
#Decomposition of Solar  - Type: Additive
ts1_renewable_data2_filtered_decompose<-decompose(ts1_renewable_data2_filtered,type="additive")
plot(ts1_renewable_data2_filtered_decompose)

#Decomposition of Wind - Type: Additive 
ts2_renewable_data2_filtered_decompose<-decompose(ts2_renewable_data2_filtered,type="additive")
plot(ts2_renewable_data2_filtered_decompose)

```


### Q5

Use the decompose function again but now change the type of the seasonal component from additive to multiplicative. What happened to the random component this time?

Answer: The random component of solar energy consumption looks random at first, but it seems to have some seasonality after that. The wind energy one shows more random pattern through 15 or 16 unit of time. After that, it also seems to have some seasonality. 

```{r}
#Decomposition of Solar - Multiplicative 
ts1_renewable_data2_filtered_decompose_M<-decompose(ts1_renewable_data2_filtered,type="multiplicative")
plot(ts1_renewable_data2_filtered_decompose_M)

#Decomposition of Wind - Multiplicative 
ts2_renewable_data2_filtered_decompose_M<-decompose(ts2_renewable_data2_filtered,type="multiplicative")
plot(ts2_renewable_data2_filtered_decompose_M)
```

### Q6

When fitting a model to this data, do you think you need all the historical data? Think about the data from 90s and early 20s. Are there any information from those years we might need to forecast the next six months of Solar and/or Wind consumption. Explain your response.

> Answer: I don't think we need all the historical data, seeing that the trend has significantly changed in more recent years for both Solar and Wind energy consumption. 

### Q7

Create a new time series object where historical data starts on January 2012. Hint: use `filter()` function so that you don't need to point to row numbers, .i.e, `filter(xxxx, year(Date) >= 2012 )`. Apply the decompose function `type=additive` to this new time series. Comment the results. Does the random component look random? Think about our discussion in class about seasonal components that depends on the level of the series. 

```{r}
#Converted to Date (Checked my code with AI)
renewable_data2_filtered$Month <- as.Date(paste0(renewable_data2_filtered$Month, "-01"), format = "%Y-%m-%d")
head(renewable_data2_filtered)

#Filtered out the historical data starts on Jan 2012
renewable_data2_filtered_recent <- renewable_data2_filtered %>%
  filter(Month >= as.Date("2012-01-01"))

#1 Time Series object starts on Jan 2012 - Solar
ts1_renewable_data2_filtered_recent <- ts(renewable_data2_filtered_recent$`Solar Energy Consumption`, 
                                          start = c(2012, 1), 
                                          frequency = 12)

#2 Time Series object starts on Jan 2012 - Wind 
ts2_renewable_data2_filtered_recent <- ts(renewable_data2_filtered_recent$`Wind Energy Consumption`, 
                                          start = c(2012, 1), 
                                          frequency = 12)

#1 Decomposition - Solar 
ts1_renewable_data2_filtered_recent_decompose<-decompose(ts1_renewable_data2_filtered_recent,type="additive")
plot(ts1_renewable_data2_filtered_recent_decompose)

#2 Decomposition - Wind 
ts2_renewable_data2_filtered_recent_decompose<-decompose(ts2_renewable_data2_filtered_recent,type="additive")
plot(ts2_renewable_data2_filtered_recent_decompose)
```

> Answer: According to the decomposed time series of Solar and Wind, the random component in more recent data appears to be more unpredictable compared to the results in Q4 and Q5. The overall trends continue to rise, which contrasts with the previous trend component that remained stable in the past before shifting upward after a certain period.


## Identify and Remove outliers

### Q8

Apply the `tsclean()` to both series from Q7. Did the function removed any outliers from the series? Hint: Use `autoplot()` to check if there is difference between cleaned series and original series.

Answer: Yes, it seems like the outliers are removed from the series. Compared to the original plots, the cleaned ones are less fluctuating with fewer outliers. 

```{r}
#1 Clean - Solar 
ts1_clean<-tsclean(ts1_renewable_data2_filtered)

#2 Clean - Wind 
ts2_clean <-tsclean(ts2_renewable_data2_filtered)

#Cleaned ver. autoplots 
autoplot(ts1_clean)
autoplot(ts2_clean)

#Original plot 
print(plot_ts1)
print(plot_ts2)
```

### Q9

Redo number Q8 but now with the time series you created on Q7, i.e., the series starting in 2014. Using what `autoplot()` again what happened now?Did the function removed any outliers from the series?



```{r}
#1 Clean - Solar 
ts1_clean_recent<-tsclean(ts1_renewable_data2_filtered_recent)

#2 Clean - Wind 
ts2_clean_recent <-tsclean(ts2_renewable_data2_filtered_recent)

#Cleaned ver. autoplots
autoplot(ts1_clean_recent)
autoplot(ts2_clean_recent)

#Original autoplots
autoplot(ts1_renewable_data2_filtered_recent)
autoplot(ts2_renewable_data2_filtered_recent)

####
```

> Answer:
Answer: Similar to Q8, the plots with tsclean() function tend to have fewer outliers than the original plots. Compared to the original plots, the cleaned plots are less fluctuating with fewer outliers. 
